Estimation of number of runaway electrons per avalanche in Earth's atmosphere

### 2020-10-03 EPL-2020.03/main.tex
EPL third revision

    A recent paper \cite{Oreshkin_2018} presents the results of the 3D Monte-Carlo simulation of runaway avalanche development. The paper is focused on the avalanche-to-streamer transition \revision{and consideration of a critical avalanche (That is, an avalanche in which the self-field becomes comparable to the external field. This condition is similar to the criterion of the avalanche-to-streamer transition for a Townsend electron avalanche~\cite{Meek1940}. The self-field is determined not only by the number of electrons, but also by the transverse size of the avalanche, therefore the transition to the critical stage is determined also by the transverse diffusion of the runaway electrons~\cite{Dwyer2010}. But in our calculation, the self-field of the avalanche is not taken into account, because the avalanche does not reach the critical stage in our simulation.)}. One of the main results of \cite{Oreshkin_2018} is the estimation of the number of electrons in one avalanche: $10^{17}$ - $10^{18}$. The value is significantly larger than values presented in other simulations~\cite{Gurevich:2001, dwyer2003fundamental,dwyer2011low}, which show about $10^6$ fast electrons and $10^{10}$ slow electrons per avalanche in normal conditions (air pressure, gas composition, etc.). The difference could be attributed to some non-standard approximations made in the article.

@@

    The simulation data was fitted by exponential function from Eq.~\ref{eq:exp} with $N_0 = 1$ and $l_a$ as a free parameter. The resulting value is $l_a \approx 85\un{m}$ without positrons, and $l_a \approx 78\un{m}$ with positrons. For comparison, Eq.~\ref{eq:dwyer}~(Dwyer's formula) predicts $l_a \approx 64\un{m}$ . Differences between our and Dwyer's result can be explained by the fact that Dwyer conducted simulation for normal condition and extrapolated results for another pressure. In addition, differences in the configuration of the Monte-Carlo model can contribute. Using received values for characteristic length, we can estimate the full number of runaway electrons in avalanches for different acceleration lengths. The results are presented in Table~\ref{tab:approx}. For reasonable acceleration cell lengths  $1200 - 1700\un{meters}$ it gives only $10^6-10^8$ runaway electrons, which is by many order less than $10^{17}$ - $10^{18}$ predicted by \cite{Oreshkin_2018}. $10^{17}$ - $10^{18}$~runaway electrons could be obtained only for cell lengths closer to $4000\un{meters}$ which are not possible in real atmospheric conditions~\cite[p.69-85]{rakov_uman}. It should be noted that we considered an avalanche created by a single seed electron, while cosmic rays are usually assumed to be a source of seed particles. But the flux of electrons of the required energies lies in the range of $10^3-10^4\un{m^{-2}s^{-1}}$~\cite{dwyer2003fundamental, sato2015analytical, sato2016analytical, dorman2013cosmic}. And taking into account that the movement of one avalanche through a cloud $1-1.5\un{km}$ length takes less than $10\un{\mu s}$ of the avalanche created by these electrons is quite separated in time and space. So that even if we take one single avalanche and a set of avalanches created with cosmic rays, then we still will not be able to achieve the value obtained by \cite{Oreshkin_2018}. \revision{Of course, we are considering various sources (for example, an extensive air showers can give a huge number of secondary electrons, however, in the context of studying thunderstorms, it is rather an exotic event~\cite{murzin_2007_book, grieder_2010_book}), but the results of simulations in ~\cite{Oreshkin_2018} are given per one seed electron, and this value is therefore comparable.}

    \begin{table}[h]
        \centering
        \begin{tabular}{crrr}
            \hline
            & & \multicolumn{2}{r}{Number of runaway electrons} \\
            &   Length, m &   without positrons &  with positrons \\
            \hline
            \multirow{4}*{\rotatebox[origin=c]{90}{ simulation}} & 300 &  34.3  &  46 \\
            & 500 &  361     &  589 \\
            & 700 &  3802     &  7539 \\
            & 800 &  12350 &  --- \\
            \hline
            \multirow{5}*{\rotatebox[origin=c]{90}{extrapolation}}& 1200 &  1.4e+06 &  4.3e+06 \\
            & 1700 &  5.0e+08 &  2.5e+09 \\
            & 2000 &  1.7e+10 &  1.2e+11 \\
            & 4000 &  2.9e+20 &  1.3e+22 \\
            & 5000 &  3.7e+25 &  4.5e+27 \\
            \hline
        \end{tabular}
        \caption{Estimation of the total number of runaway electrons based on $700-800\un{meter}$ simulation. The first part of the table is taken from the simulation, the second part is the extrapolation of simulation data.}
        \label{tab:approx}
    \end{table}

    When discussing light\revision{n}ing breakdown, it is important to compute not only number of breakdown electrons, but also the total number of electrons and ions generated. This number could be very roughly estimated from energy conservation law.

>>>>

    A recent paper \cite{Oreshkin_2018} presents the results of the 3D Monte-Carlo simulation of runaway avalanche development. The paper is focused on the avalanche-to-streamer transition and consideration of a critical avalanche (That is, an avalanche in which the self-field becomes comparable to the external field. This condition is similar to the criterion of the avalanche-to-streamer transition for a Townsend electron avalanche~\cite{Meek1940}. The self-field is determined not only by the number of electrons, but also by the transverse size of the avalanche, therefore the transition to the critical stage is determined also by the transverse diffusion of the runaway electrons~\cite{Dwyer2010}. But in our calculation, the self-field of the avalanche is not taken into account, because the avalanche does not reach the critical stage in our simulation.). One of the main results of \cite{Oreshkin_2018} is the estimation of the number of electrons in one avalanche: $10^{17}$ - $10^{18}$. The value is significantly larger than values presented in other simulations~\cite{Gurevich:2001, dwyer2003fundamental,dwyer2011low}, which show about $10^6$ fast electrons and $10^{10}$ slow electrons per avalanche in normal conditions (air pressure, gas composition, etc.). The difference could be attributed to some non-standard approximations made in the article.

@@

    The simulation data was fitted by exponential function from Eq.~\ref{eq:exp} with $N_0 = 1$ and $l_a$ as a free parameter. The resulting value is $l_a \approx 85\un{m}$ without positrons, and $l_a \approx 78\un{m}$ with positrons. For comparison, Eq.~\ref{eq:dwyer}~(Dwyer's formula) predicts $l_a \approx 64\un{m}$ . Differences between our and Dwyer's result can be explained by the fact that Dwyer conducted simulation for normal condition and extrapolated results for another pressure. In addition, differences in the configuration of the Monte-Carlo model can contribute. Using received values for characteristic length, we can estimate the full number of runaway electrons in avalanches for different acceleration lengths. The results are presented in Table~\ref{tab:approx}. For reasonable acceleration cell lengths  $1200 - 1700\un{meters}$ it gives only $10^6-10^8$ runaway electrons, which is by many order less than $10^{17}$ - $10^{18}$ predicted by \cite{Oreshkin_2018}. $10^{17}$ - $10^{18}$~runaway electrons could be obtained only for cell lengths closer to $4000\un{meters}$ which are not possible in real atmospheric conditions~\cite[p.69-85]{rakov_uman}. It should be noted that we considered an avalanche created by a single seed electron, while cosmic rays are usually assumed to be a source of seed particles. But the flux of electrons of the required energies lies in the range of $10^3-10^4\un{m^{-2}s^{-1}}$~\cite{dwyer2003fundamental, sato2015analytical, sato2016analytical, dorman2013cosmic}. And taking into account that the movement of one avalanche through a cloud $1-1.5\un{km}$ length takes less than $10\un{\mu s}$ of the avalanche created by these electrons is quite separated in time and space. So that even if we take one single avalanche and a set of avalanches created with cosmic rays, then we still will not be able to achieve the value obtained by \cite{Oreshkin_2018}. \revision{We are considering various sources: external (for example, extensive air showers can give a huge number of secondary electrons, however, in the context of studying thunderstorms, this event is rather unlikely~\cite{murzin_2007_book, grieder_2010_book}), internal (avalanche development with thermal runaway is studied in~\cite{moss2006}). Although, the results of simulations in ~\cite{Oreshkin_2018} are given per one seed electron, thus the comparison of results is independent on the number of seed electrons.} 
    \begin{table}[h]
        \centering
        \begin{tabular}{crrr}
            \hline
            & & \multicolumn{2}{r}{Number of runaway electrons} \\
            &   Length, m &   without positrons &  with positrons \\
            \hline
            \multirow{4}*{\rotatebox[origin=c]{90}{ simulation}} & 300 &  34.3  &  46 \\
            & 500 &  361     &  589 \\
            & 700 &  3802     &  7539 \\
            & 800 &  12350 &  --- \\
            \hline
            \multirow{5}*{\rotatebox[origin=c]{90}{extrapolation}}& 1200 &  1.4e+06 &  4.3e+06 \\
            & 1700 &  5.0e+08 &  2.5e+09 \\
            & 2000 &  1.7e+10 &  1.2e+11 \\
            & 4000 &  2.9e+20 &  1.3e+22 \\
            & 5000 &  3.7e+25 &  4.5e+27 \\
            \hline
        \end{tabular}
        \caption{Estimation of the total number of runaway electrons based on $700-800\un{meter}$ simulation. The first part of the table is taken from the simulation, the second part is the extrapolation of simulation data.}
        \label{tab:approx}
    \end{table}

    When discussing lightning breakdown, it is important to compute not only number of breakdown electrons, but also the total number of electrons and ions generated. This number could be very roughly estimated from energy conservation law.
